### 2024-08-23 README.Rmd
los

---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<p align="center">
  <img src="./images.jpg" width="750" title="hover text">
</p>

# NativeAdvertising

This repository contains the data and code for our paper:

> Trbusic, D., Males, D., & Sikic, L. (YYYY). Native advertising as a content marketing strategy. Name of journal/book. <

Our pre-print is online here:

> Trbusic, D., Males, D., & Sikic, L. (YYYY). Native advertising as a content marketing strategy. Name of journal/book, Accessed [Date]. Online at https://doi.org/xxx/xxx. <

### How to cite

Please cite this compendium as:

> Trbusic, D., Males, D., & Sikic, L. (2019). Compendium of R code and data for 'Native advertising as a content marketing strategy'. Accessed [Date]. Online at https://doi.org/xxx/xxx.<

### How to download or install

You can download the compendium as a zip from from this URL: </archive/master.zip>

Or you can install this compendium as an R package, rrcompendium, from GitHub with:
### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.