### 2024-08-27 analysis/paper/paper.Rmd
los

---
title: Native advertising as a content marketing strategy
author:
  - name: Davor Trbusic
    email: davor.trbusic@unicath.hr
    affiliation: Croatian Catholic University
    correspondingauthor: true
    footnote: 1
  - name: Drazen Males
    email: dmales@fhs.hr
    affiliation: Faculty of Croatian Studies
  - name: Luka Sikic
    email: luka.sikic@unicath.hr
    affiliation: Croatian Catholic University
    footnote: 2
address:
  - code: Croatian Catholic University
    organization: Croatian Catholic University
    addressline: Ilica 242
    city: Zagreb
    state: 
    postcode: 10000
    country: Croatia
  - code: Faculty of Croatian Studies
    organization: Faculty of Croatian Studies
    addressline: Borongajska cesta 83d
    postcode: 10000
    city: Zagreb
    country: Croatia
footnote:
  - code: 1
    text: "This is the corresponding author."
  - code: 2
    text: "Co-author of the study."
abstract: |
  This research focuses on analyzing the content elements of native advertisements on the most popular Croatian online portals, with the aim of identifying key characteristics of headlines and content, as well as their correlation with audience reach. The study was conducted on a sample of 543 native ads published on six leading Croatian online portals from December 2021 to May 2022. The content analysis method was used, focusing on two general categories: headline characteristics and content characteristics. The analysis found that nearly 80% of the headlines had at least one clickbait characteristic, with the most commonly used styles including uncertainty (44.38%), use of numbers (9.39%), and emphasis on emotions (5.89%). The headlines were predominantly declarative sentences, with one-fifth being interrogative sentences, while the connection between the ad headline and the advertiser's brand name was very rarely present. In terms of content, visual elements were present in all ads, with photographs being the dominant element, and visual identity elements of the advertiser, such as logos, were included in 34.44% of the ads. Regarding sources in native ads, 47.33% of the ads did not use statements, while the most common sources were individuals from the organization (19.71%). Statements from direct users of the products or services were present in 6.81% of the cases, while statements from brand ambassadors and influencers were present in smaller percentages (4.79% and 3.5%, respectively). These findings highlight the dominance of clickbait headlines and the importance of visual identity in native advertising, while direct brand association and the use of various sources, including direct users, appear less frequently. A comparison of different types of headlines and sources with ad reach indicates specific practices in native advertising across different industries.
keywords: 
  - native advertising
  - content analysis
  - headline characteristics
journal: "An awesome journal"
date: "`r Sys.Date()`"
linenumbers: false
numbersections: true
bibliography: refs.bib
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: kable

---

# Introduction

The concentration of so-called hybrid forms of media content in the contemporary online environment is becoming increasingly intense (Balasubramanian, 1994; Macnamara, 2014; Taiminen, Luoma-aho & Tolvanen, 2015). Among the most prominent forms are sponsored content (Tutaj & Reijmersdal, 2012), content marketing (Pulizzi, 2014; Pulizzi & Piper, 2023), brand journalism (Cole & Greer, 2013), and native advertising (Verčič & Tkalac Verčič, 2016). For most researchers, the latter term will serve as an umbrella term that attempts to simplify the classification of hybrid forms, as native advertising by definition implies a paid advertisement that resembles editorial content, aiming to attract the target audience with useful content while simultaneously promoting the brand, values, and products (Cole & Greer, 2013; Howe & Teufel, 2014; Wojdynski & Evans, 2016). On the other hand, the research by Taiminen et al. (2015) showed that most public relations and marketing professionals preferred to use the term “content marketing” when discussing commercial hybrid forms of online content. In addition to the obvious terminological confusion, three other problematic areas of native advertising need to be highlighted:

> Application area: Is native advertising a public relations technique, a marketing tool, or a tactic that does not differ from traditional advertising in any way?

> Authorship of native ads: Who creatively signs them and who is responsible for their content?

> Legal and ethical framework: How is native advertising legally defined, and what is the practice of media houses in labeling such advertisements (Trbušić, Maleš & Šikić, 2024).

The theoretical basis of this paper is the thesis that native advertising is “an extension of content marketing and provides marketers with a platform to place content in front of a wider audience they wouldn't ordinarily reach” (Crook, 2022). Although there are authors who highlight certain criteria by which content marketing differs from native advertising, which will be described in more detail later in the paper, the ultimate goal is common: to drive profitable customer action. The profitability of such market communication for organizations and brands is demonstrated by recent metrics. Figures show that 47% of marketing professionals globally believe in the effectiveness of this type of advertising (Kloot, 2022), and that in 2023, spending on native advertising in the United States increased by 12% compared to the previous year (eMarketer, 2023). Thus, with more than 97 billion US dollars invested, native advertising holds the largest share (59.7%) of the total expenditure for all types of advertising (eMarketer, 2023).

<details>
<summary>Click to expand the code chunk</summary>

```{r warning=FALSE, message=FALSE}

library(dplyr)
library(readr)
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(readxl)
library(xlsx)
library(here)
library(knitr)
library(kableExtra)

```

</details>

# Native advertising and content marketing – two faces of the same coin?
Charmaine Du Plessis (2015a) defines content marketing as sharing brand content via owned media to gain earned media, which clearly implies that it is not a type of media content that is paid for, as (native) advertising certainly is. In this context, Du Plessis employs the classic PESO model of media content in public relations (Dietrich, 2014), which distinguishes paid media (advertising or other sponsored content); earned media (publicity generated from media pitches or news releases); shared media (content shared and re-shared on social media); and owned media (such as company’s website, newsletter, or email communications). The order in the PESO model has long reflected the priorities of organizations in media strategy planning. However, recent research (Macnamara, Lwin, Adi & Zerfass, 2016) indicates a shift in the order in practice, with the dominance shifting to the SOEP model instead of the classic PESO model. The reasons for this are manifold, but the most indicative is that owned and earned media content are trusted more than paid media, while shared media (i.e., “recommendations from people I know” and “consumer opinions posted online”) are the most trusted information sources of all (Nielsen, 2015).

Like Du Plessis, other authors (Handley & Chapman, 2011; Halvorson & Rach, 2012) also differentiate content marketing from other forms of hybrid content exclusively based on the distribution channel (owned media). However, there are authors who leave much more room for interpretation, not strictly confining it to owned media that later generates earned media (publicity). Their definitions are more focused on the strategy, intent, and ultimate goal of content marketing, rather than on the tactics and type of media content within the PESO framework. For instance, they highlight that content marketing is “a strategic marketing approach focused on creating and distributing valuable, relevant, and consistent content” (Pulizzi & Piper, 2023; CMI, 2024), which is disseminated on various platforms such as newspapers, magazines, TV, or radio channels (Holliman & Rowley, 2014).

The distribution of valuable, relevant, and consistent content is also at the core of native advertising (Campbell & Marks, 2015; Kim, 2017), which is especially evident in so-called branded or native content. According to the Interactive Advertising Bureau (IAB), there are three fundamental forms of native advertising: In-Feed/In-Content native advertising, Content Recommendation Ads, and Branded/native content. The latter refers to “paid content by a brand that is published in the same format as full-fledged editorial content on the publisher's website, usually in collaboration with the publisher's own teams” (IAB, 2019, p. 14). This form of native advertising will be the focus of the research part of this paper for two main reasons: it conceptually correlates with the forms and tactics of content marketing, and it is the only form among the mentioned types of native advertising that can be approached analytically.

Types and tactics of content marketing typically include blogs, infographics, video content, photos, webinars, podcasts and social media (Halligan & Shah, 2010; Handley & Chapman, 2011; Bloomstein, 2012; Wuebben, 2012; Vibrant Publishers, 2020), but also paid ads (Baker, 2024), such as sponsored advertising options on social media, affiliate marketing, remarketing, native ads etc. (Jefferson & Tanton, 2015). It is yet another proof of the inconsistent interpretation of content marketing as something that is disseminated exclusively via owned media. It is more appropriate to consider content marketing as a set of tactics that, among other things, can drive traffic back to the website of some company or product (owned media). It is also noticeable that some authors of definitions avoid identifying content marketing with product advertising which aims at the purchase of a product or service (Patrutiu Baltes, 2015), seeking to further differentiate it from other related marketing tactics. The fact is that content marketing is part of an inbound marketing strategy, which, unlike the outbound strategy, is focused on “attracting valuable consumers (potential, existing, or aspirational) who choose to interact with a particular company that provides them with something useful” (Opreana & Vinerean, 2015, p. 29). However, it is misguided to claim that the ultimate goal of any marketing communication, including content marketing, is not sales or gaining profit (Rowley, 2008; Pulizzi & Piper, 2023). Furthermore, both academics and practitioners agree that content marketing relies on a consumer-centric strategy (Ho, Pang & Choy, 2020). In conclusion, regardless of the type of marketing communication, whether it is content marketing, native advertising, or brand journalism - although one can discuss nuanced tactics and approaches to attract consumers into the marketing funnel and later motivate them for retention and advocacy phases - profitable customer action remains the intrinsic motive of every organization or brand.

Previous research does not encompass the analysis of native advertising as part of content marketing; rather, these concepts have been the focus of separate studies. Thus, insights in the context of native advertising relate to the analysis of readers' cognitive perception of native ads and their persuasive elements (Wojdynski & Evans, 2016), analysis of communication strategies in native ads (Wang & Li, 2017), perception of the organization or the sponsor of native ads and its socially responsible activities (Jung & Heo, 2018; Krouwer, Poels & Paulussen, 2019; Beckert, Koch, Viererbl, Denner, & Peter, 2020; Wu & Overton, 2021), perception of the credibility of online portals that publish native ads (Howe & Teufel, 2014), and analysis of the basic features of native ads based on the most-read selected Croatian online portals (Trbušić et al., 2024). Research in the field of content marketing primarily reflects valuable theoretical contributions, such as conceptual definitions, classification of types and forms of content marketing, identification of strategies and tactics of content marketing in the digital environment (Koiso-Kanttila, 2004; Rowley, 2008; Handley & Chapman, 2011; Bloomstein, 2012; Wuebben, 2012; Pulizzi, 2014; Jefferson & Tanton, 2015; Vinerean, 2017; Pulizzi & Piper, 2023), and the definition of six fundamental elements of content marketing which should be considered by marketers (Du Plessis, 2015a; Du Plessis, 2015b). Empirically, content marketing has been analyzed as part of B2B strategies on digital channels (Holliman & Rowley, 2014) or in the context of content creation practices in successful companies (Du Plessis, 2015c; Ho et al., 2020). Additionally, studies have addressed the causal linkage between content marketing and online consumer behavior (Du Plessis, 2022) and content marketing as part of business strategy in the luxury industry (Rios, 2016).

# Methodology and analytical design

The goal of the research is to determine the basic characteristics of the content elements of native ads (as one of the strategies of content marketing) on the observed most read Croatian internet portals. The specific objectives of the research are focused on the observation of two broader entities. It is about the features of the title (e.g. types of sentences in the titles, representation and types of clickbait titles, mention of the brand name) and the analysis of elements within the content of native ads (e.g. citing sources, i.e. use of interlocutor statements, use of equipment and visual identity). The mentioned categories were not chosen arbitrarily, but were defined according to the types and typology of content marketing (see p. 5). In addition, the goal is to determine whether there is a significant correlation between the characteristics of the title and content of native ads with audience reach. Equally, attention is also focused in the work on the specifics that are observed in this type of advertising in different industries, i.e. different practices resorted to by those ordering native ads.

<details>
<summary>Click to expand the code chunk</summary>

```{r warning=FALSE, message=FALSE}

original <- read.xlsx(here("analysis", "data", "raw_data", "native_articles.xlsx"), sheetIndex = 1) %>%
  mutate(V1 = as.numeric(V1))

variables <- read.xlsx(here("analysis", "data", "raw_data", "native_research.xlsx"), sheetIndex = 1)

dta <- merge(original %>% filter(V1 %in% variables$V1),
  variables %>% select(-c(23, 24, 25)),
  by = "V1",
  all.x = TRUE)

# Convert the 'DATE' column to Date format
dta$DATE <- as.Date(dta$DATE)

stemmed <- readRDS("C:/Users/Lukas/Dropbox/Mediatoolkit/native_token_stemm.rds")
stemmed <- readRDS("C:/Users/lukas/Dropbox/Članci/Native rad/native_token_stemm.rds")

stemmed <- stemmed %>%
  rename("wordp" = "word") %>%
  rename("word" = "transformed_column")

```
</details>

# REZULTATI ZA KOMBINIRANJE

### summarizes and visualizes the distribution of articles by industry

<details>
<summary>Click to expand the code chunk</summary>
```{r warning=FALSE, message=FALSE,error=FALSE}

industry_list <- data.frame(
  id = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28),
  category = c(
    "Financijska industrija",
    "Građevinska industrija",
    "Moda i ljepota",
    "Politička institucija",
    "Sportska industrija",
    "Zdravstvena industrija",
    "Obrazovanje i znanost",
    "Tehnologija",
    "Auto-moto industrija",
    "ICT industrija",
    "Kultura i umjetnost",
    "Turizam, odmor, putovanja, ugostiteljstvo",
    "Energetska industrija",
    "Prehrambena industrija",
    "Industrija bezalkoholnih pića",
    "Alkoholna pića",
    "Industrija nekretnina",
    "Maloprodaja",
    "Igre na sreću",
    "Medijske kuće i marketinške, digitalne agencije",
    "Namještaj, saloni namještaja, kućanske potrepštine, kućanski uređaji, sredstva za čišćenje",
    "Trgovački / shopping centri",
    "Glazbena industrija",
    "Filmska industrija",
    "Brodogradnja",
    "Pošta / distribucija pošiljki",
    "Komunalne usluge"
  )
)

IND <- dta %>%
  group_by(INDUSTRIJA) %>%
  summarise(N = n()) %>% 
  mutate(percentIND = round(N / sum(N) * 100,2))

# Create a histogram of the 'price' variable in the 'diamonds' dataset
# ggplot(data = IND, aes(x = percentIND)) +
#   geom_histogram(binwidth = 1, fill = "blue", alpha = 0.7) +
#   ggtitle("Industrija") +
#   xlab("Industrija") +
#   ylab("%") +
#   scale_x_continuous(breaks = seq(from = min(dta$INDUSTRIJA), to =  max(dta$INDUSTRIJA), by = 1))

# Joining the dataframes
IND_joined <- IND %>%
  left_join(industry_list, by = c("INDUSTRIJA" = "id")) %>%
  arrange(desc(N))

# Reordering the levels of 'category'
IND_joined$category <- with(IND_joined, reorder(category, percentIND))

#Plotting
ggplot(data = IND_joined, aes(x = category, y = percentIND)) +
  geom_bar(stat = "identity", fill = "black") +
  labs(title = "Industrija", x = "Industrija", y = "%") +
  theme_minimal() +
  # theme(
  #   axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
  #   panel.grid.major.x = element_blank(),  # Remove vertical grid lines
  #   panel.border = element_rect(fill = NA, color = "black")  # Add border
  # ) +
  coord_flip()  # Flip coordinates to have horizontal bars

# IND %>%
#   datatable(., options = list(scrollX = TRUE, scrollY = "500px"))
# 
# 
# industry_list %>%
#   datatable(., options = list(scrollX = TRUE, scrollY = "500px"))

IND_joined <- IND %>%
  left_join(industry_list, by = c("INDUSTRIJA" = "id"))

IND_final <- IND_joined %>%
  select(category, N, percentIND) %>%
  arrange(desc(N))

IND_kable <- IND_final %>%
  kable(format = "html", 
        col.names = c("Industrija", "Broj objava", "%"),
        caption = "Pregled po industrijama") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)

IND_kable

```

</details>

```{r echo=F, eval=T, message=F , warning= FALSE}
IND_kable
```

### summarizes the total reach per portal, provides descriptive statistics for reach, and creates a histogram of reach values (filtered to be between 0 and 30,000)

```{r echo=F, eval=T, message=F , warning= FALSE}
# Portals by reach
reach <- dta %>%
  group_by(FROM) %>%
  summarise(reach = sum(REACH)) %>%
  arrange(desc(reach))

reach

# make detailed descriptive statistics

summary(dta$REACH)

filtered_dta <- dta %>% filter(REACH >= 0 & REACH <= 30000)

# Create the histogram
hist(filtered_dta$REACH, 
     xlab = "Reach", 
     ylab = "Frequency", 
     main = "Reach of Portals", 
     col = "blue", 
     breaks = 50)

```

### summary statistics for reach (mean, median, min, max, total). It then connects these statistics with the corresponding clickbait descriptions

```{r echo=F, eval=T, message=F , warning= FALSE}

# Create the clickbait DataFrame
clickbait <- data.frame(
  number = c(1, 2, 3, 4, 5, 6, 7),
  text = c("Nedefinirane zamjenice", "Neizvjesnost", "Stil obrnutog narativa",
           "Naglašavanje emocija", "Korištenje brojkama", "Više kombinacija",
           "Nije clickbait naslov")
)

# Assuming 'dta' is your data frame and it contains 'REACH' and 'CLICKBAIT' columns
# Group by CLICKBAIT and summarize REACH statistics
summary_stats <- dta %>%
  group_by(CLICKBAIT) %>%
  summarise(
    count = n(),
    mean_reach = mean(REACH, na.rm = TRUE),
    median_reach = median(REACH, na.rm = TRUE),
    min_reach = min(REACH, na.rm = TRUE),
    max_reach = max(REACH, na.rm = TRUE),
    total_reach = sum(REACH, na.rm = TRUE)
  )

# Connect with clickbait names
result <- summary_stats %>%
  left_join(clickbait, by = c("CLICKBAIT" = "number"))

# Display the result
result %>%
  select(text, count, mean_reach, median_reach, min_reach, max_reach, total_reach) %>%
  kable(format = "html", 
        col.names = c("Clickbait Type", "Count", "Mean Reach", "Median Reach", 
                      "Min Reach", "Max Reach", "Total Reach"),
        caption = "Summary Statistics by Clickbait Type") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, position = "left")

```

### calculates summary statistics for reach, and then merges these results with the descriptions of each source type

```{r echo=F, eval=T, message=F , warning= FALSE}
# Create the 'izvori' DataFrame
izvori <- data.frame(
  number = c(1, 2, 3, 4, 5, 6, 7),
  text = c(
    "Nema izvora / izjave",
    "Izvor iz organizacije / tvrtke",
    "Brend ambasador",
    "Influencer, celebrity, slavna / poznata osoba",
    "Korisnik proizvoda ili usluge (poduzetnik, Marica iz Benkovca)",
    "Više izvora",
    "Izvor nepovezan s proizvodom ili uslugom koja se promovira"
  )
)

# Group by CLICKBAIT and summarize REACH statistics
summary_stats <- dta %>%
  group_by(`IZVORI...IZJAVE`) %>%
  summarise(
    count = n(),
    mean_reach = mean(REACH, na.rm = TRUE),
    median_reach = median(REACH, na.rm = TRUE),
    min_reach = min(REACH, na.rm = TRUE),
    max_reach = max(REACH, na.rm = TRUE),
    total_reach = sum(REACH, na.rm = TRUE)
  )

# Connect with clickbait names
result <- summary_stats %>%
  left_join(izvori, by = c(`IZVORI...IZJAVE` = "number"))

# Display the result
result %>%
  select(text, count, mean_reach, median_reach, min_reach, max_reach, total_reach) %>%
  kable(format = "html", 
        col.names = c("Source Type", "Count", "Mean Reach", "Median Reach", 
                      "Min Reach", "Max Reach", "Total Reach"),
        caption = "Summary Statistics by Source Type") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, position = "left")

```

### correlation between different clickbait types across industries/correlation matrix to examine relationships between clickbait usage in various industries.

```{r  echo=F, eval=T, message=F , warning= FALSE}

# Summarize the data by both INDUSTRIJA and CLICKBAIT
summary_data <- dta %>%
  group_by(INDUSTRIJA, CLICKBAIT) %>%
  summarise(N = n()) %>%
  ungroup() %>%
  mutate(percent = round(N / sum(N) * 100, 2))

# Merge with industry names
summary_with_industry <- summary_data %>%
  left_join(industry_list, by = c("INDUSTRIJA" = "id"))

# Merge with clickbait descriptions
final_summary <- summary_with_industry %>%
  left_join(clickbait, by = c("CLICKBAIT" = "number"))

# View the final table
# final_summary %>%
#   select(category, text, N, percent) %>%
#   kable(format = "html", 
#         col.names = c("Industry", "Clickbait Type", "Count", "Percentage"),
#         caption = "Distribution of Clickbait Types Across Industries") %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
#                 full_width = F, position = "left")
# Optionally, calculate correlations between industries and clickbait types
# You could reshape the data to a wide format and then calculate correlations if needed

# Example: Reshape data to wide format for correlation analysis
wide_summary <- final_summary %>%
  select(category, text, N) %>%
  spread(key = text, value = N, fill = 0)

# Calculate correlations
correlations <- cor(wide_summary[-1], use = "pairwise.complete.obs")

#print(correlations)

correlation_df <- as.data.frame(round(correlations, 2))

# Display the correlation matrix as a nicely formatted HTML table
correlation_df %>%
  rownames_to_column(var = "Clickbait Type") %>%
  kable(format = "html", 
        caption = "Correlation Matrix of Clickbait Types Across Industries") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, position = "left")

```

### correlation between different source types across industries/correlation matrix to explore how different source types are used together within various industrie

```{r echo=F, eval=T, message=F , warning= FALSE}

# Step 1: Group by Industry and Source Type
industry_sources <- dta %>%
  group_by(INDUSTRIJA, `IZVORI...IZJAVE`) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  left_join(industry_list, by = c("INDUSTRIJA" = "id")) %>%
  left_join(izvori, by = c("IZVORI...IZJAVE" = "number"))

# Step 2: Reshape Data to Wide Format
wide_sources <- industry_sources %>%
  select(category, text, Count) %>%
  spread(key = text, value = Count, fill = 0)

# Step 3: Calculate the Correlation Matrix
correlation_matrix <- cor(wide_sources[-1], use = "pairwise.complete.obs")

# View the correlation matrix
#print(correlation_matrix)

correlation_df <- as.data.frame(round(correlations, 2))

# Display the correlation matrix as a nicely formatted HTML table
correlation_df %>%
  rownames_to_column(var = "Source Type") %>%
  kable(format = "html", 
        caption = "Correlation Matrix of Source Types Across Industries") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, position = "left")

```

### how different clickbait types are distributed across various industries

```{r fig.height=12, fig.width=10, echo=F, eval=T, message=F , warning= FALSE}
industry_clickbait <- dta %>%
  group_by(INDUSTRIJA, CLICKBAIT) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  left_join(industry_list, by = c("INDUSTRIJA" = "id")) %>%
  left_join(clickbait, by = c("CLICKBAIT" = "number"))

# Step 2: Calculate Percentages within each Industry
industry_clickbait <- industry_clickbait %>%
  group_by(INDUSTRIJA) %>%
  mutate(Percent = round(Count / sum(Count) * 100, 2)) %>%
  ungroup()

# Step 3: Create a Table Showing Distribution
# You can simply view the table
# industry_clickbait %>%
#   select(category, text, Count, Percent) %>%
#   kable(format = "html", 
#         col.names = c("Industry", "Clickbait Type", "Count", "Percentage"),
#         caption = "Distribution of Clickbait Types Across Industries") %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
#                 full_width = F, position = "left")
# Optional Step: Create a Heatmap for Visualization
ggplot(industry_clickbait, aes(x = category, y = text, fill = Percent)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Distribution of Clickbait Types Across Industries",
       x = "Industry",
       y = "Clickbait Type",
       fill = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### how different source types are distributed across various industries

```{r fig.height=12, fig.width=10, echo=F, eval=T, message=F , warning= FALSE}

# Assuming 'dta' is your data frame with 'INDUSTRIJA' and 'IZVORI...IZJAVE' columns

# Step 1: Group by Industry and Source Type
industry_sources <- dta %>%
  group_by(INDUSTRIJA, `IZVORI...IZJAVE`) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  left_join(industry_list, by = c("INDUSTRIJA" = "id")) %>%
  left_join(izvori, by = c("IZVORI...IZJAVE" = "number"))

# Step 2: Calculate Percentages within each Industry
industry_sources <- industry_sources %>%
  group_by(INDUSTRIJA) %>%
  mutate(Percent = round(Count / sum(Count) * 100, 2)) %>%
  ungroup()

# Step 3: Create a Table Showing Distribution
# You can simply view the table
#print(industry_sources)

# Optional Step: Create a Heatmap for Visualization
ggplot(industry_sources, aes(x = category, y = text, fill = Percent)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Distribution of Source Types Across Industries",
       x = "Industry",
       y = "Source Type",
       fill = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

The sample for the research consists of native ads published over a six-month period (December 2021 – May 2022) on the six most-read Croatian internet portals (index.hr, jutarnji.hr, vecernji.hr, 24sata.hr, telegram.hr, slobodnadalmacija.hr), which were selected as the most influential online media in the Croatian digital media space based on their traffic (Peruško, 2023). After conducting a machine-based pre-analysis that included all posts on Croatian internet portals during the observed period, totaling over 1.5 million posts across 5,104 websites, a representative sample was selected based on traffic criteria, i.e., the largest Croatian internet portals where a machine search was conducted using keywords and phrases in the content or title. The keywords used to identify native content included terms such as "paid content," "sponsored content," "native," "content provided by," "content created by," "this section was made by," "powered by," "tnative," "24contenthaus," considering that native content should be marked as such in the media. After isolating all articles over the six-month period that matched these keywords, a total of 811, those that did not meet the characteristics of native articles were excluded through human analysis, primarily referring to sponsored or PR releases and other content not related to native ads. The final sample for the research consisted of 543 native ads on six internet portals that stand out for their influence and readership in the Republic of Croatia. Additionally, before determining the research sample, it was confirmed that all internet portals whose articles were subject to content analysis offer native advertising services, through a review of their publicly available marketing, advertising, and public relations service price lists.

Out of the total sample of 543 native ads, 177 posts, or 32.6% of the total, were published on the internet portal jutarnji.hr, 103 posts, or 18.97%, on vecernji.hr, and 99 posts, or 18.23%, on telegram.hr. The internet portal 24sata.hr accounted for 88 analyzed articles, or 16.21%, index.hr for 62 articles, or 11.42%, and slobodnadalmacija.hr for 14 native articles, or 2.58%.

The study, conducted on a sample of 543 native ads published between December 1, 2021, and May 30, 2022, on six selected Croatian internet portals that meet the traffic criteria, was carried out using the content analysis method. An analytical matrix containing five general categories (general identification elements, content emphasis, headline characteristics, content characteristics, audience engagement) was applied, within which 19 variables were identified to observe each of the 543 ads. In the analysis, the authors independently read and analyzed all identified ads and then compared the results, achieving full agreement on all variables among the authors.

For the purposes of this study, two general categories were observed, in accordance with the types and typology of content marketing: headline characteristics and content characteristics of native ads. The variables related to the categorization of headline types, the connection of the headline with the advertiser's brand, and the determination of the type of clickbait headline, while the analysis of the native article content focused on the type of visual content (photo, video, infographic, graphic, combinations of these) and its connection with the visual identity of the advertiser, as well as the representation of different types of sources/statements (source from the organization/company, brand ambassador, influencer - celebrity, direct user of the product or service, multiple sources) in the native ads.

The aim of the study was also to determine whether there is a significant correlation between the types of headlines and audience reach. Therefore, data on headline characteristics were correlated with data on the reach of each individual ad. Additionally, the study examined whether certain types of native ad headlines were more prevalent or characteristic of different industries. Similarly, the types of statements found in the analyzed native ads were compared and observed in relation to the type of industry to which the advertiser belongs and the reach of the ad.

# Results and Discussion

From the total number of 543 analyzed native ads that were published in a six-month period (December 2021 - May 2022) on six Croatian internet portals that meet the criteria of attendance and relevance (jutarnji.hr, index.hr, vecernji.hr, 24sata.hr) , telegram.hr, slobodandalmacija.hr), almost four-fifths of them (79.01%) were found to have clickbait titles. 
When determining the characteristics of clickbait, we used the typology defined by Kanižaj... The research showed that clickbait titles characterized by uncertainty (44.38%) are the most represented. Multiple combinations of styles in titles were observed in 13.81% of ads, while 20.99% of titles were without a single clickbait feature. 
Titles in which the style that implies the use of numbers prevails appear in 9.39% of cases, and 5.89% of the titles are characterized by emphasizing emotions. The use of undefined pronouns was observed in 4.97% of the titles, while the reverse narrative style was registered in 0.55% of the ads.

<!-- <details> -->
<!-- <summary>Click to expand the code chunk</summary> -->

```{r echo=F, eval=T, message=F , warning= FALSE}

# Create the clickbait DataFrame
clickbait <- data.frame(
  number = c(1, 2, 3, 4, 5, 6, 7),
  text = c("Nedefinirane zamjenice", "Neizvjesnost", "Stil obrnutog narativa",
           "Naglašavanje emocija", "Korištenje brojkama", "Više kombinacija",
           "Nije clickbait naslov")
)

# Grouping, summarizing, and calculating percentages
result <- dta %>%
  group_by(CLICKBAIT) %>%
  summarise(N = n()) %>%
  mutate(percent = round(N / sum(N) * 100, 2)) %>%
  arrange(desc(N)) %>%
  left_join(clickbait, by = c("CLICKBAIT" = "number"))

# Create a beautiful table
result %>%
  select(text, N, percent) %>%
  kable("html", col.names = c("Description", "Count", "%")) %>%
  add_header_above(c("Table 1. Prevalence of clickbait headline characteristics in the analyzed native ads" = 3)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE, position = "left") %>%
  column_spec(1, width = "20em") %>%
  column_spec(2, width = "10em") %>%
  column_spec(3, width = "10em")

```
<!-- </details> -->

In the vast majority of the analyzed native ads (94.48%), the advertiser's brand name was not mentioned in the headline, while in 5.52%, or thirty out of the 543 articles, the brand name was included in the headline. Here, we will add the distinction between a sponsored article, which typically mentions the advertiser in the headline, and native advertising, which generally does not, as it employs a subtler advertising tactic.

The majority of the headlines, 71.09%, were declarative sentences, while one-fifth (20.26%) were interrogative, and 8.95% were exclamatory sentences. Regarding the content characteristics of the observed native ads, it is noteworthy that all analyzed ads included visual elements, with photography being the most prominent, either alone or in combination with other types of visual content. As shown in Table 2, in 73.67% of cases, the ads were equipped exclusively with photographic content, while in 11.97% of cases, the ads included both photographic and graphic content. In 11.05% of ads, photographic content was accompanied by video content, and in 2.39% of cases, the native ads contained a combination of photo, video, and graphic content.

<!-- <details> -->
<!-- <summary>Click to expand the code chunk</summary> -->

```{r echo=F, eval=T, message=F , warning= FALSE}

# Create the 'oprema' DataFrame
oprema <- data.frame(
  number = c(1, 2, 3, 4, 5, 6, 7, 8),
  text = c(
    "Foto-sadržaj",
    "Video-sadržaj",
    "Grafika ili infografika (ne uključuje logo)",
    "Foto i video sadržaj",
    "Foto i grafički sadržaj",
    "Video i grafički sadržaj",
    "Foto, video i grafički sadržaj",
    "Članak nema nikakvu opremu"
  )
)

# Grouping, summarizing, and calculating percentages
result <- dta %>%
  group_by(OPREMA) %>%
  summarise(N = n()) %>%
  mutate(percent = round(N / sum(N) * 100, 2)) %>%
  arrange(desc(N)) %>%
  left_join(oprema, by = c("OPREMA" = "number"))

# Create a beautiful table
result %>%
  select(text, N, percent) %>%
  kable("html", col.names = c("Description", "Count", "%")) %>%
  add_header_above(c("Table 2. Type of visual equipment" = 3)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE, position = "left") %>%
  column_spec(1, width = "30em") %>%
  column_spec(2, width = "10em") %>%
  column_spec(3, width = "10em")

```
<!-- </details> -->

The advertiser's logo, or elements of the advertiser's visual identity, were featured as part of the visual content in 34.44% of the analyzed native ads. However, in 356 out of the 543 analyzed articles, this was not the case, meaning they did not contain a visual link to the advertiser.

Among the content characteristics of native ads in this study, the types of sources or statements within the articles were also examined, as shown in Table 3. The ratio of ads that use sources to those that do not is approximately similar. Among the ads that include a source, the most common type is a person from the organization or the advertiser (19.71%). The second most prominent, with 10.31% representation, are sources not directly connected to the product or service, while in 7.55% of cases, multiple sources were present. In 6.81% of ads, the source was a testimony or statement from someone presented as a direct user of the product or service. Well-known individuals are also used as sources in native ads; brand ambassadors were used in 4.79% of cases, while influencers or celebrities were featured in 3.5% of the observed ads.

<!-- <details> -->
<!-- <summary>Click to expand the code chunk</summary> -->

```{r echo=F, eval=T, message=F , warning= FALSE}

# Create the 'izvori' DataFrame
izvori <- data.frame(
  number = c(1, 2, 3, 4, 5, 6, 7),
  text = c(
    "Nema izvora / izjave",
    "Izvor iz organizacije / tvrtke",
    "Brend ambasador",
    "Influencer, celebrity, slavna / poznata osoba",
    "Korisnik proizvoda ili usluge (poduzetnik, Marica iz Benkovca)",
    "Više izvora",
    "Izvor nepovezan s proizvodom ili uslugom koja se promovira"
  )
)

# Grouping, summarizing, and calculating percentages
result <- dta %>%
  group_by(`IZVORI...IZJAVE`) %>%
  summarise(N = n()) %>%
  mutate(percent = round(N / sum(N) * 100, 2)) %>%
  arrange(desc(N)) %>%
  left_join(izvori, by = c("IZVORI...IZJAVE" = "number"))

# Create a beautiful table
result %>%
  select(text, N, percent) %>%
  kable("html", col.names = c("Description", "Count", "%")) %>%
  add_header_above(c("Table 3. Representation of different types of sources in the statements" = 3)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE, position = "left") %>%
  column_spec(1, width = "30em") %>%
  column_spec(2, width = "10em") %>%
  column_spec(3, width = "10em")

```
<!-- </details> -->

# Conclusion

The conclusion of the research can be summarized through several key points. An analysis of 543 native ads published on six Croatian internet portals showed a significant presence of clickbait titles, where almost 80% of the titles had at least one clickbait characteristic, with the suspense style being most often used. Although a smaller part of the headlines (5.52%) mentioned the brand, most of them were neutral about direct advertising. Titles are mostly declarative sentences, while visual elements, especially photos, were present in all ads, with the advertiser's logo included in 34.44% of cases. In the area of ​​sources of statements, almost half of the articles did not use statements, while in other articles the sources were most often people from the organization, with a relatively small presence of influencers and celebrities. These findings indicate the dominance of clickbait titles and the significant role of visual equipment in native advertising, while direct association with the brand and the use of authoritative sources are less common.

# References