### 2024-09-10 analysis/paper/paper.Rmd
los

---
title: Native advertising as a content marketing strategy
author:
  - name: Davor Trbusic
    email: davor.trbusic@unicath.hr
    affiliation: Croatian Catholic University
    correspondingauthor: true
    footnote: 1
  - name: Drazen Males
    email: dmales@fhs.hr
    affiliation: Faculty of Croatian Studies
  - name: Luka Sikic
    email: luka.sikic@unicath.hr
    affiliation: Croatian Catholic University
    footnote: 2
address:
  - code: Croatian Catholic University
    organization: Croatian Catholic University
    addressline: Ilica 242
    city: Zagreb
    state: 
    postcode: 10000
    country: Croatia
  - code: Faculty of Croatian Studies
    organization: Faculty of Croatian Studies
    addressline: Borongajska cesta 83d
    postcode: 10000
    city: Zagreb
    country: Croatia
footnote:
  - code: 1
    text: "This is the corresponding author."
  - code: 2
    text: "Co-author of the study."
abstract: |
  This research focuses on analyzing the content elements of native advertisements on the most popular Croatian online portals, with the aim of identifying key characteristics of headlines and content, as well as their correlation with audience reach. The study was conducted on a sample of 543 native ads published on six leading Croatian online portals from December 2021 to May 2022. The content analysis method was used, focusing on two general categories: headline characteristics and content characteristics. The analysis found that nearly 80% of the headlines had at least one clickbait characteristic, with the most commonly used styles including uncertainty (44.38%), use of numbers (9.39%), and emphasis on emotions (5.89%). The headlines were predominantly declarative sentences, with one-fifth being interrogative sentences, while the connection between the ad headline and the advertiser's brand name was very rarely present. In terms of content, visual elements were present in all ads, with photographs being the dominant element, and visual identity elements of the advertiser, such as logos, were included in 34.44% of the ads. Regarding sources in native ads, 47.33% of the ads did not use statements, while the most common sources were individuals from the organization (19.71%). Statements from direct users of the products or services were present in 6.81% of the cases, while statements from brand ambassadors and influencers were present in smaller percentages (4.79% and 3.5%, respectively). These findings highlight the dominance of clickbait headlines and the importance of visual identity in native advertising, while direct brand association and the use of various sources, including direct users, appear less frequently. A comparison of different types of headlines and sources with ad reach indicates specific practices in native advertising across different industries.
keywords: 
  - native advertising
  - content analysis
  - headline characteristics
journal: "An awesome journal"
date: "`r Sys.Date()`"
linenumbers: false
numbersections: true
bibliography: refs.bib
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: kable

---

# Introduction

The concentration of so-called hybrid forms of media content in the contemporary online environment is becoming increasingly intense [@balasubramanian1994; @macnamara2014; @taiminen2015]. Among the most prominent forms are sponsored content [@tutaj2012], content marketing [@pulizzi2014; @pulizzi2023], brand journalism [@cole2013], and native advertising [@vercic2016]. For most researchers, the latter term will serve as an umbrella term that attempts to simplify the classification of hybrid forms, as native advertising by definition implies a paid advertisement that resembles editorial content, aiming to attract the target audience with useful content while simultaneously promoting the brand, values, and products [@cole2013; @howe2014; @wojdynski2016]. On the other hand, the research by [@taiminen2015] showed that most public relations and marketing professionals preferred to use the term “content marketing” when discussing commercial hybrid forms of online content. In addition to the obvious terminological confusion, three other problematic areas of native advertising need to be highlighted:

>	Application area: Is native advertising a public relations technique, a marketing tool, or a tactic that does not differ from traditional advertising in any way?

>	Authorship of native ads: Who creatively signs them and who is responsible for their content?

>	Legal and ethical framework: How is native advertising legally defined, and what is the practice of media houses in labelling such advertisements [@trbusic2024]?

The theoretical basis of this paper is the thesis that native advertising is “an extension of content marketing and provides marketers with a platform to place content in front of a wider audience they wouldn't ordinarily reach” [@crook2022]. Although there are authors who highlight certain criteria by which content marketing differs from native advertising, which will be described in more detail later in the paper, the ultimate goal is common: to drive profitable customer action. The profitability of such market communication for organizations and brands is demonstrated by recent metrics. Figures show that 47% of marketing professionals globally believe in the effectiveness of this type of advertising [@kloot2022], and that in 2023, spending on native advertising in the United States increased by 12% compared to the previous year [@emarketer2023]. Thus, with more than 97 billion US dollars invested, native advertising holds the largest share (59.7%) of the total expenditure for all types of advertising [@emarketer2023].

# Native advertising and content marketing – two faces of the same coin?

[@duplessis2015a] defines content marketing as sharing brand content via owned media to gain earned media, which clearly implies that it is not a type of media content that is paid for, as (native) advertising certainly is. In this context, Du Plessis employs the classic PESO model of media content in public relations [@dietrich2014], which distinguishes paid media (advertising or other sponsored content); earned media (publicity generated from media pitches or news releases); shared media (content shared and re-shared on social media); and owned media (such as company’s website, newsletter, or email communications). The order in the PESO model has long reflected the priorities of organizations in media strategy planning. However, recent research [@macnamara2016] indicates a shift in the order in practice, with the dominance shifting to the SOEP model instead of the classic PESO model. The reasons for this are manifold, but the most indicative is that owned and earned media content are trusted more than paid media, while shared media (i.e., “recommendations from people I know” and “consumer opinions posted online”) are the most trusted information sources of all [@nielsen2015].
Like Du Plessis, other authors [@handley2011; @halvorson2012] also differentiate content marketing from other forms of hybrid content exclusively based on the distribution channel (owned media). However, there are authors who leave much more room for interpretation, not strictly confining it to owned media that later generates earned media (publicity). Their definitions are more focused on the strategy, intent, and ultimate goal of content marketing, rather than on the tactics and type of media content within the PESO framework. For instance, they highlight that content marketing is “a strategic marketing approach focused on creating and distributing valuable, relevant, and consistent content” [@pulizzi2023; @cmi2023], which is disseminated on various platforms such as newspapers, magazines, TV, or radio channels [@holliman2014].

The distribution of valuable, relevant, and consistent content is also at the core of native advertising [@campbell2015; @kim2017], which is especially evident in so-called branded or native content. According to the Interactive Advertising Bureau (IAB), there are three fundamental forms of native advertising: In-Feed/In-Content native advertising, Content Recommendation Ads, and Branded/native content. The latter refers to “paid content by a brand that is published in the same format as full-fledged editorial content on the publisher's website, usually in collaboration with the publisher's own teams” [@iab2019]. This form of native advertising will be the focus of the research part of this paper for two main reasons: it conceptually correlates with the forms and tactics of content marketing, and it is the only form among the mentioned types of native advertising that can be approached analytically.

Types and tactics of content marketing typically include blogs, infographics, video content, photos, webinars, podcasts and social media [@halligan2010; @handley2011; @bloomstein2012; @wuebben2011; @vibrant2020], but also paid ads [@baker2024], such as sponsored advertising options on social media, affiliate marketing, remarketing, native ads etc. [@jefferson2015]. It is yet another proof of the inconsistent interpretation of content marketing as something that is disseminated exclusively via owned media. It is more appropriate to consider content marketing as a set of tactics that, among other things, can drive traffic back to the website of some company or product (owned media). It is also noticeable that some authors of definitions avoid identifying content marketing with product advertising which aims at the purchase of a product or service [@patrutiu2015], seeking to further differentiate it from other related marketing tactics. The fact is that content marketing is part of an inbound marketing strategy, which, unlike the outbound strategy, is focused on “attracting valuable consumers (potential, existing, or aspirational) who choose to interact with a particular company that provides them with something useful” [@opreana2015]. However, it is misguided to claim that the ultimate goal of any marketing communication, including content marketing, is not sales or gaining profit [@rowley2008; @pulizzi2023]. Furthermore, both academics and practitioners agree that content marketing relies on a consumer-centric strategy [@ho2020]. In conclusion, regardless of the type of marketing communication, whether it is content marketing, native advertising, or brand journalism - although one can discuss nuanced tactics and approaches to attract consumers into the marketing funnel and later motivate them for retention and advocacy phases - profitable customer action remains the intrinsic motive of every organization or brand.

Previous research does not encompass the analysis of native advertising as part of content marketing; rather, these concepts have been the focus of separate studies. Thus, insights in the context of native advertising relate to the analysis of readers' cognitive perception of native ads and their persuasive elements [@wojdynski2016], analysis of communication strategies in native ads [@wang2017], perception of the organization or the sponsor of native ads and its socially responsible activities [@jung2018; @krouwer2020; @beckert2020; @wu2021], perception of the credibility of online portals that publish native ads [@howe2014], and analysis of the basic features of native ads based on the most-read selected Croatian online portals [@trbusic2024]. Research in the field of content marketing primarily reflects valuable theoretical contributions, such as conceptual definitions, classification of types and forms of content marketing, identification of strategies and tactics of content marketing in the digital environment [@koiso2004; @rowley2008; @handley2011; @bloomstein2012; @wuebben2011; @pulizzi2014; @jefferson2015; @vinerean2017; @pulizzi2023], and the definition of six fundamental elements of content marketing which should be considered by marketers [@duplessis2015a; @duplessis2015b]. Empirically, content marketing has been analysed as part of B2B strategies on digital channels [@holliman2014] or in the context of content creation practices in successful companies [@duplessis2015c; @ho2020]. Additionally, studies have addressed the causal linkage between content marketing and online consumer behaviour [@duplessis2022] and content marketing as part of business strategy in the luxury industry [@rios2016].

# Methodology and analytical design

The objective of this study is to determine the content characteristics of native advertisements (as one of the content marketing strategies) on the most-read Croatian internet portals. Additionally, the study aims to establish whether there is a relationship between the characteristics of native ad headlines and content with audience reach. Attention is also directed towards the specific practices observed in different industries, particularly the diverse approaches employed by sponsors of native advertisements.

<details>
<summary>Import libraries/click to see the code</summary>

```{r warning=FALSE, message=FALSE}

library(dplyr)
library(readr)
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(readxl)
library(xlsx)
library(here)
library(knitr)
library(kableExtra)

```

</details>

The data used in the study consists of native advertisements published over a six-month period (December 2021 – May 2022) on the six most-read Croatian internet portals (index.hr, jutarnji.hr, vecernji.hr, 24sata.hr, telegram.hr, slobodnadalmacija.hr) [@perusko2023]. The data was sampled based on the combination of machine and human filtering of the population of posts on Croatian internet portals during the observed period, totalling over 1.5 million posts on 5,104 web pages. The machine search was conducted based on selected keywords and phrases (“paid content”, “sponsored content”, “native”, “content provided by”, “content created by”, “powered by” and “tnative”) in the content or headlines of the article. This procedure resulted in 811 articles. After the thorough human filtering, irrelevant articles were manually excluded and the final empirical sample consisted of 543 native articles. Prior to determining the research sample, it was confirmed that all analysed internet portals offer native advertising services through a review of their publicly available marketing, advertising, and public relations pricing lists.

<details>
<summary>Import data/click to see the code</summary>

```{r warning=FALSE, message=FALSE}

original <- read.xlsx(here("analysis", "data", "raw_data", "native_articles.xlsx"), sheetIndex = 1) %>%
  mutate(V1 = as.numeric(V1))

variables <- read.xlsx(here("analysis", "data", "raw_data", "native_research.xlsx"), sheetIndex = 1)

dta <- merge(original %>% filter(V1 %in% variables$V1),
  variables %>% select(-c(23, 24, 25)),
  by = "V1",
  all.x = TRUE)

# Convert the 'DATE' column to Date format
dta$DATE <- as.Date(dta$DATE)

stemmed <- readRDS("C:/Users/Lukas/Dropbox/Mediatoolkit/native_token_stemm.rds")
stemmed <- readRDS("C:/Users/lukas/Dropbox/Članci/Native rad/native_token_stemm.rds")

stemmed <- stemmed %>%
  rename("wordp" = "word") %>%
  rename("word" = "transformed_column")

```
</details>

Of the total sample of 543 native advertisements, 177 posts, or 32.6%, were from the portal jutarnji.hr, 103 posts or 18.97% from vecernji.hr, and 99 posts or 18.23% from telegram.hr. The portal 24sata.hr accounted for 88 analysed articles or 16.21%, index.hr for 62 or 11.42%, and slobodnadalmacija.hr for 14 native articles or 2.58%.

The research on a sample of 543 native advertisements published between December 1, 2021, and May 30, 2022, on the six selected Croatian internet portals, was conducted using content analysis methodology. An analytical matrix was applied, containing five general categories (general identification elements, content emphasis, headline characteristics, content characteristics, audience engagement), within which 19 variables were observed for each of the 543 ads. The authors independently read and analysed all identified ads and then compared results, achieving full agreement across all variables.

For the purposes of this study, two general categories were selected for the analysis, in line with content marketing types and typologies: headline characteristics and content characteristics of native advertisements. The focus was on categorizing headline types, their connection with the advertiser's brand, and identifying the type of clickbait headline. The content analysis of native articles focused on the type of visual equipment (photo, video, infographic, graphics, combinations of these) and its relationship with the advertiser's visual identity, as well as the representation of different types of sources/statements (source from the organization/company, brand ambassador, influencer/celebrity, direct user of the product or service, multiple sources) in native ads. The study also examined whether certain types of native ad headlines were more prevalent or characteristic of different industries. Similarly, the types of statements found in the analysed native ads were compared and observed concerning the industry type and the reach of each ad.

# Reults and discussion

Of the total 543 analysed native advertisements published over a six-month period (December 2021 – May 2022) on six Croatian internet portals that meet the criteria of visit frequency and relevance (jutarnji.hr, index.hr, vecernji.hr, 24sata.hr, telegram.hr, slobodnadalmacija.hr), nearly four-fifths (79.01%) were found to contain clickbait headlines.

<details>
<summary>Some descriptive statistics/click to see the code</summary>
```{r warning=FALSE, message=FALSE,error=FALSE, echo=FALSE, results='hide'}

industry_list <- data.frame(
  id = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28),
  category = c(
    "Financijska industrija",
    "Građevinska industrija",
    "Moda i ljepota",
    "Politička institucija",
    "Sportska industrija",
    "Zdravstvena industrija",
    "Obrazovanje i znanost",
    "Tehnologija",
    "Auto-moto industrija",
    "ICT industrija",
    "Kultura i umjetnost",
    "Turizam, odmor, putovanja, ugostiteljstvo",
    "Energetska industrija",
    "Prehrambena industrija",
    "Industrija bezalkoholnih pića",
    "Alkoholna pića",
    "Industrija nekretnina",
    "Maloprodaja",
    "Igre na sreću",
    "Medijske kuće i marketinške, digitalne agencije",
    "Namještaj, saloni namještaja, kućanske potrepštine, kućanski uređaji, sredstva za čišćenje",
    "Trgovački / shopping centri",
    "Glazbena industrija",
    "Filmska industrija",
    "Brodogradnja",
    "Pošta / distribucija pošiljki",
    "Komunalne usluge"
  )
)

IND <- dta %>%
  group_by(INDUSTRIJA) %>%
  summarise(N = n()) %>% 
  mutate(percentIND = round(N / sum(N) * 100,2))

# Create a histogram of the 'price' variable in the 'diamonds' dataset
# ggplot(data = IND, aes(x = percentIND)) +
#   geom_histogram(binwidth = 1, fill = "blue", alpha = 0.7) +
#   ggtitle("Industrija") +
#   xlab("Industrija") +
#   ylab("%") +
#   scale_x_continuous(breaks = seq(from = min(dta$INDUSTRIJA), to =  max(dta$INDUSTRIJA), by = 1))

# Joining the dataframes
IND_joined <- IND %>%
  left_join(industry_list, by = c("INDUSTRIJA" = "id")) %>%
  arrange(desc(N))

# Reordering the levels of 'category'
IND_joined$category <- with(IND_joined, reorder(category, percentIND))

#Plotting
ggplot(data = IND_joined, aes(x = category, y = percentIND)) +
  geom_bar(stat = "identity", fill = "black") +
  labs(title = "Industrija", x = "Industrija", y = "%") +
  theme_minimal() +
  # theme(
  #   axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
  #   panel.grid.major.x = element_blank(),  # Remove vertical grid lines
  #   panel.border = element_rect(fill = NA, color = "black")  # Add border
  # ) +
  coord_flip()  # Flip coordinates to have horizontal bars

# IND %>%
#   datatable(., options = list(scrollX = TRUE, scrollY = "500px"))
# 
# 
# industry_list %>%
#   datatable(., options = list(scrollX = TRUE, scrollY = "500px"))

IND_joined <- IND %>%
  left_join(industry_list, by = c("INDUSTRIJA" = "id"))

IND_final <- IND_joined %>%
  select(category, N, percentIND) %>%
  arrange(desc(N))

IND_kable <- IND_final %>%
  kable(format = "html", 
        col.names = c("Industrija", "Broj objava", "%"),
        caption = "Pregled po industrijama") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)

IND_kable

```

```{r echo=F, eval=T, message=F , warning= FALSE, results='hide'}
IND_kable
```

```{r echo=F, eval=T, message=F , warning= FALSE, results='hide'}
# Portals by reach
reach <- dta %>%
  group_by(FROM) %>%
  summarise(reach = sum(REACH)) %>%
  arrange(desc(reach))

reach

# make detailed descriptive statistics

summary(dta$REACH)

filtered_dta <- dta %>% filter(REACH >= 0 & REACH <= 30000)

# Create the histogram
hist(filtered_dta$REACH, 
     xlab = "Reach", 
     ylab = "Frequency", 
     main = "Reach of Portals", 
     col = "blue", 
     breaks = 50)

```

</details>

In determining clickbait headlines, the categorization provided by [@chen2015] was used: vague pronouns, uncertainty, reverse narrative style, emphasis on emotions, and the use of numbers. The study revealed that the most prevalent clickbait headlines were those characterized by uncertainty (44.38%). Multiple combinations of styles within headlines were observed in 13.81% of the ads, while 20.99% of the headlines exhibited no clickbait characteristics at all. Headlines dominated by numerical content appeared in 9.39% of cases, while 5.89% of headlines emphasized emotional appeal. Headlines using vague pronouns were noted in 4.97% of cases, and the reverse narrative style was registered in 0.55% of the ads.

```{r echo=F, eval=T, message=F , warning= FALSE}

# Create the clickbait DataFrame
clickbait <- data.frame(
  number = c(1, 2, 3, 4, 5, 6, 7),
  text = c("Use of vague pronouns", "Uncertainty", "Reverse narrative style",
           "Emphasis on emotions", "Use of numbers", "Multiple combinations",
           "Not a clickbait headline")
)

# Assuming 'dta' is your data frame and it contains 'REACH' and 'CLICKBAIT' columns
# Group by CLICKBAIT and summarize REACH statistics
summary_stats <- dta %>%
  group_by(CLICKBAIT) %>%
  summarise(
    count = n(),
    mean_reach = mean(REACH, na.rm = TRUE),
    median_reach = median(REACH, na.rm = TRUE),
    min_reach = min(REACH, na.rm = TRUE),
    max_reach = max(REACH, na.rm = TRUE),
    total_reach = sum(REACH, na.rm = TRUE)
  )

# Connect with clickbait names
result <- summary_stats %>%
  left_join(clickbait, by = c("CLICKBAIT" = "number"))

# Display the result
result %>%
  select(text, count, mean_reach, median_reach, min_reach, max_reach, total_reach) %>%
  kable(format = "html", 
        col.names = c("Clickbait Type", "Count", "Mean Reach", "Median Reach", 
                      "Min Reach", "Max Reach", "Total Reach"),
        caption = "Table 1: Prevalence and reach of clickbait headline characteristics in the analysed native ads (n=543)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, position = "left")

```

In the further analysis of data displayed in Table 1, various clickbait headline techniques were examined to determine their effectiveness in driving audience engagement. The data provide insights into how different strategies influence the reach of advertisements, emphasizing the role of headlines as a primary trigger for audience interaction.

The use of uncertainty as a clickbait technique in native adverts yields an average reach of 11,202.718, indicating a generally high level of engagement compared to other analysed categories. The median reach for this category is 10,051, suggesting that at least half of these ads successfully engage over 10,000 people. The minimum reach recorded is 963, highlighting that even the least successful ads using uncertainty headlines still manage to attract a considerable audience. The maximum reach is an impressive 142,818, demonstrating that some ads in this category have much higher reach compared to other types of clickbait headlines. The total reach across all adverts employing uncertainty is 2,699,855, making it the category with the most substantial overall impact.

For ads with non-clickbait headlines, the mean reach is 8,990.412, which is lower compared to most clickbait categories. The median reach stands at 8,400, indicating consistent but moderate engagement. The minimum reach in this category is 963, like the uncertainty category, reflecting a baseline level of engagement. The maximum reach is 32,270, which is significantly lower than the top clickbait categories, suggesting that non-clickbait headlines are less likely to go viral. The total reach for this category is 1,024,907, indicating a more modest overall impact compared to clickbait strategies.

Native ads using multiple combinations of clickbait techniques achieve the highest mean reach at 12,504.333, demonstrating strong effectiveness. The median reach is 10,891, showing consistently high engagement across this category. The minimum reach is 963, consistent with other categories, ensuring a same baseline. The maximum reach is 80,240, indicating a significant potential for very high engagement when multiple techniques are combined. The total reach for this category is 937,825, which is notable given the smaller number of ads utilizing this approach.

The technique of incorporating numbers in native article headlines results in a mean reach of 10,610.373, indicating that this strategy is quite effective. The median reach is 10,418, suggesting consistent performance. The minimum reach is 2,129, reflecting a higher baseline reach compared to some other categories. The maximum reach is 18,790, indicating moderate potential for high engagement. The total reach for this category is 541,129, indicating a substantial, though not overwhelming, overall impact.

Ads that emphasize emotions have the lowest mean reach among the categories at 7,367.219, indicating lesser effectiveness. The median reach is 6,125, showing that these ads generally reach a smaller audience. The minimum reach is 1,531, which is slightly higher than the minimum reach in some other categories, but still relatively low. The maximum reach is 19,962, suggesting limited potential for virality. The total reach for this category is 235,751, the lowest total reach among all categories, reflecting a more limited overall impact.

The use of vague pronouns in headlines is one of the more effective techniques, with a mean reach of 11,954.889. The median reach is 10,701, indicating consistently strong engagement. The minimum reach is 5,688, showing a solid baseline performance. The maximum reach is 24,296, indicating some potential for high engagement, though not as much as seen in the uncertainty or multiple combinations categories. The total reach for this category is 322,782, which is relatively strong considering the fewer ads in this category.

Finally, as displayed in Table 1, native adverts employing a reverse narrative style have the lowest mean reach at 6,861.667, indicating limited effectiveness. The median reach is 5,688, showing that this technique generally attracts a smaller audience. Interestingly, the minimum reach is also 5,688, matching the median, which suggests very consistent performance, albeit at a lower level. The maximum reach is 9,209, the lowest maximum reach across all categories, reinforcing the limited appeal of this style. The total reach is just 20,585, the lowest total reach, reflecting minimal overall impact.

This comprehensive analysis provides a clear view of the effectiveness of different clickbait headline techniques in engaging audiences. Techniques such as uncertainty and multiple combinations of clickbait elements stand out as particularly effective, while approaches like reverse narrative style and emphasis on emotions are less impactful in terms of reach.

In most of the analysed native ads (94.48%), the advertiser's brand name was not mentioned in the headline, while in 5.52% of cases, or in thirty out of the 543 articles, the brand name was included in the headline. Most headlines, 71.09%, were declarative sentences, while one-fifth of the headlines (20.26%) were interrogative, and 8.95% were exclamatory sentences. In terms of content characteristics of the observed native ads, it is noteworthy that all analysed ads featured visual content, with photographs being the most prominent, either alone or in combination with other types of visual content. As shown in Table 2, 73.67% of the ads were equipped exclusively with photographic content, while in 11.97% of cases, both photographs and graphics were included. In 11.05% of the ads, photographs were accompanied by video content, and in 2.39% of cases, the native ads contained a combination of photo, video, and graphics.

```{r echo=F, eval=T, message=F , warning= FALSE}

# Create the 'oprema' DataFrame
oprema <- data.frame(
  number = c(1, 2, 3, 4, 5, 6, 7, 8),
  text = c(
    "Photo content",
    "Video content",
    "Graphics or infographics",
    "Photo and video content",
    "Photo and graphics content",
    "Video and graphics content",
    "Photo, video, and graphics content",
    "Article lacks visual elements"
  )
)

# Grouping, summarizing, and calculating percentages
result <- dta %>%
  group_by(OPREMA) %>%
  summarise(N = n()) %>%
  mutate(percent = round(N / sum(N) * 100, 2)) %>%
  arrange(desc(N)) %>%
  left_join(oprema, by = c("OPREMA" = "number"))

# Create a beautiful table
result %>%
  select(text, N, percent) %>%
  kable("html", col.names = c("Description", "Count", "%")) %>%
  add_header_above(c("Table 2. Type of visual equipment" = 3)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE, position = "left") %>%
  column_spec(1, width = "30em") %>%
  column_spec(2, width = "10em") %>%
  column_spec(3, width = "10em")

```

The advertiser's logo, or elements of the advertiser's visual identity, were included in the visual content of the analysed native ads in 34.44% of cases, while in 356 out of 543 analysed articles, this was not the case, meaning they did not contain a visual link to the advertiser.

Regarding the content characteristics of native ads in this study, the types of sources or statements in the articles were also observed, as shown in Table 3. The ratio of ads using sources to those that do not is approximately similar, with the most common sources in the ads being individuals from the organization, i.e., the advertiser (19.71%). The second most common sources, with a prevalence of 10.31%, were those not directly connected to the product or service, while multiple sources were present in 7.55% of the ads, and in 6.81% of the ads, there was testimony or a statement from a person presented as a direct user of the product or service. Celebrities were also used as sources in native ads, with brand ambassadors’ statements used in 4.79% of cases, and influencers or celebrities were present in 3.5% of the analysed ads.

Table 3. Representation of different types of sources in the statements and reach (n=543)

```{r echo=F, eval=T, message=F , warning= FALSE}
# Create the 'izvori' DataFrame
izvori <- data.frame(
  number = c(1, 2, 3, 4, 5, 6, 7),
  text = c(
    "No source / statement",
    "Source from the organization / company",
    "Brand ambassador",
    "Influencer / celebrity",
    "Direct user of the product or service",
    "Multiple sources",
    "Source unrelated to the product or service"
  )
)

# Group by CLICKBAIT and summarize REACH statistics
summary_stats <- dta %>%
  group_by(`IZVORI...IZJAVE`) %>%
  summarise(
    count = n(),
    mean_reach = mean(REACH, na.rm = TRUE),
    median_reach = median(REACH, na.rm = TRUE),
    min_reach = min(REACH, na.rm = TRUE),
    max_reach = max(REACH, na.rm = TRUE),
    total_reach = sum(REACH, na.rm = TRUE)
  )

# Connect with clickbait names
result <- summary_stats %>%
  left_join(izvori, by = c(`IZVORI...IZJAVE` = "number"))

# Display the result
result %>%
  select(text, count, mean_reach, median_reach, min_reach, max_reach, total_reach) %>%
  kable(format = "html", 
        col.names = c("Source Type", "Count", "Mean Reach", "Median Reach", 
                      "Min Reach", "Max Reach", "Total Reach"),
        caption = "Table 3. Representation of different types of sources in the statements and reach (n=543)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, position = "left")

```

In the analysis of Table 3, various types of sources were examined in relation to their impact on the reach of native ads. The findings reveal distinct differences in the effectiveness of advertising strategies based on the source type. Native ads featuring a statement from brand ambassador have an average reach of 10,057.192, similar in effectiveness to ads without a source. The median reach is 9,719, indicating consistent performance just below the mean. The minimum reach is 1,558, reflecting a decent baseline level of engagement. The maximum reach is 31,308 and the total reach is 261,487, which, while lower, still represents a significant impact given the smaller number of ads in this category.

As final of the categories, native ads featuring a statement given by influencer or celebrity achieve one of the highest average reaches at 14,645.789, indicating strong effectiveness compared to other types of sources. The median reach is 14,797, suggesting very consistent and high engagement. The minimum reach is 9,975, the highest minimum reach across all categories, showing that these ads generally perform well. The maximum reach is 21,300, indicating strong potential for high engagement, though not as high as ads using multiple sources. The total reach is 278,270, which is impressive given the relatively small number of ads, highlighting the high effectiveness of this strategy.

```{r  echo=F, eval=T, message=F , warning= FALSE, results='hide'}

# Summarize the data by both INDUSTRIJA and CLICKBAIT
summary_data <- dta %>%
  group_by(INDUSTRIJA, CLICKBAIT) %>%
  summarise(N = n()) %>%
  ungroup() %>%
  mutate(percent = round(N / sum(N) * 100, 2))

# Merge with industry names
summary_with_industry <- summary_data %>%
  left_join(industry_list, by = c("INDUSTRIJA" = "id"))

# Merge with clickbait descriptions
final_summary <- summary_with_industry %>%
  left_join(clickbait, by = c("CLICKBAIT" = "number"))

# View the final table
# final_summary %>%
#   select(category, text, N, percent) %>%
#   kable(format = "html", 
#         col.names = c("Industry", "Clickbait Type", "Count", "Percentage"),
#         caption = "Distribution of Clickbait Types Across Industries") %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
#                 full_width = F, position = "left")
# Optionally, calculate correlations between industries and clickbait types
# You could reshape the data to a wide format and then calculate correlations if needed

# Example: Reshape data to wide format for correlation analysis
wide_summary <- final_summary %>%
  select(category, text, N) %>%
  spread(key = text, value = N, fill = 0)

# Calculate correlations
correlations <- cor(wide_summary[-1], use = "pairwise.complete.obs")

#print(correlations)

correlation_df <- as.data.frame(round(correlations, 2))

# Display the correlation matrix as a nicely formatted HTML table
correlation_df %>%
  rownames_to_column(var = "Clickbait Type") %>%
  kable(format = "html", 
        caption = "Correlation Matrix of Clickbait Types Across Industries") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, position = "left")

```

```{r echo=F, eval=T, message=F , warning= FALSE, results='hide'}

# Step 1: Group by Industry and Source Type
industry_sources <- dta %>%
  group_by(INDUSTRIJA, `IZVORI...IZJAVE`) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  left_join(industry_list, by = c("INDUSTRIJA" = "id")) %>%
  left_join(izvori, by = c("IZVORI...IZJAVE" = "number"))

# Step 2: Reshape Data to Wide Format
wide_sources <- industry_sources %>%
  select(category, text, Count) %>%
  spread(key = text, value = Count, fill = 0)

# Step 3: Calculate the Correlation Matrix
correlation_matrix <- cor(wide_sources[-1], use = "pairwise.complete.obs")

# View the correlation matrix
#print(correlation_matrix)

correlation_df <- as.data.frame(round(correlations, 2))

# Display the correlation matrix as a nicely formatted HTML table
correlation_df %>%
  rownames_to_column(var = "Source Type") %>%
  kable(format = "html", 
        caption = "Correlation Matrix of Source Types Across Industries") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, position = "left")

```

```{r fig.height=12, fig.width=10, echo=F, eval=T, message=F , warning= FALSE}
industry_clickbait <- dta %>%
  group_by(INDUSTRIJA, CLICKBAIT) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  left_join(industry_list, by = c("INDUSTRIJA" = "id")) %>%
  left_join(clickbait, by = c("CLICKBAIT" = "number"))

# Step 2: Calculate Percentages within each Industry
industry_clickbait <- industry_clickbait %>%
  group_by(INDUSTRIJA) %>%
  mutate(Percent = round(Count / sum(Count) * 100, 2)) %>%
  ungroup()

# Step 3: Create a Table Showing Distribution
# You can simply view the table
# industry_clickbait %>%
#   select(category, text, Count, Percent) %>%
#   kable(format = "html", 
#         col.names = c("Industry", "Clickbait Type", "Count", "Percentage"),
#         caption = "Distribution of Clickbait Types Across Industries") %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
#                 full_width = F, position = "left")
# Optional Step: Create a Heatmap for Visualization
ggplot(industry_clickbait, aes(x = category, y = text, fill = Percent)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Graph 1: Distribution of different clickbait types across various industries",
       x = "Industry",
       y = "Clickbait Type",
       fill = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The histogram titled “Distribution of Clickbait Types Across Industries” (Histogram 1) illustrates the prevalence of various clickbait headline techniques across a range of industries. The chart visually represents how different sectors employ specific strategies to attract audience engagement, with colour intensity indicating the percentage of ads that utilize each technique within each industry (darkest shade of blue indicate maximal prevalence of each technique within the industry, while lighter shades close to purple and all the way to white, indicate lower intensity – as shown in legend).

In general, the data reveal that most industries adopt a mix of clickbait strategies, tailored to the nature of their content and target audience. However, distinct patterns emerge, showing that certain sectors exhibit a clear preference for specific techniques. For instance, industries dealing with uncertainty, especially in contemporary times, such as financial services, healthcare and energetics sector, tend to favour headlines that evoke uncertainty to engage readers, which is also most widely used technique across industries.

Conversely, sectors like shipbuilding prefer straightforward, non-clickbait headlines, likely reflecting the technical and factual nature of their content, while this technique is also widely used in fashion industry, as the data displayed in Histogram 1 is showing. When it comes to the specific types of clickbait, the use of multiple combinations of techniques is particularly prominent in the movie industry along with fashion, and arts & culture sector, indicating this strategic approach to maximizing engagement. The reverse narrative style, though generally less commonly used, finds notable application in the sports industry, suggesting that reversing the expected narrative could be effective in this sector. This technique is also used in minimal amounts in retail and ICT industries, while the other industries tend not to apply this strategy in their native advertising.

Vague pronouns is also lesser used technique, and it is more favoured approach in media, marketing and digital agencies, along with postal services. In contrast, emotional emphasis is more frequently used in the utility services sector, science and education and alcoholic beverages industry, where emotionally charged messaging can influence consumer behaviour. The use of numbers in headlines is moderately preferred in the retail industries, along with furniture and home appliances industry, and shopping centres where presenting clear, quantifiable information can appeal to a detail-oriented audience.

```{r fig.height=12, fig.width=10, echo=F, eval=T, message=F , warning= FALSE}

# Step 1: Group by Industry and Source Type
industry_sources <- dta %>%
  group_by(INDUSTRIJA, `IZVORI...IZJAVE`) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  left_join(industry_list, by = c("INDUSTRIJA" = "id")) %>%
  left_join(izvori, by = c("IZVORI...IZJAVE" = "number"))

# Step 2: Calculate Percentages within each Industry
industry_sources <- industry_sources %>%
  group_by(INDUSTRIJA) %>%
  mutate(Percent = round(Count / sum(Count) * 100, 2)) %>%
  ungroup()

# Step 3: Create a Table Showing Distribution
# You can simply view the table
#print(industry_sources)

# Optional Step: Create a Heatmap for Visualization
ggplot(industry_sources, aes(x = category, y = text, fill = Percent)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Distribution of Source Types Across Industries",
       x = "Industry",
       y = "Source Type",
       fill = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

The histogram titled “Distribution of Source Types Across Industries” (Histogram 2) illustrates how different types of sources are distributed across various industries in the context of advertising or content creation. The vertical axis lists the source types, while the horizontal axis represents different industries.

In general, the histogram shows that industries tend to favour certain source types depending on the nature of the industry and its target audience. For example, the “No Source or Statement” category is prominently used across multiple industries, particularly in the movies industry, utility services sector and shopping centres, with maximum prevalence in those areas. The same category, which implies lacking any sources in native adverts is also widely used in automotive industries, along with retail, furniture and home appliances tourism, and postal services. This technique is significantly present in all of industries except shipbuilding, real estate and arts & culture sector, with no articles without statement in this analysis. The use of sources from within the organization or company is second most widely employed technique, and it is most common in sectors like shipbuilding and real estate, where insider knowledge and authority are crucial for establishing trust with the audience. Fashion and beauty industry also widely use this approach, along with arts & culture and ICT sectors, and in political institutions.

The use of multiple sources is widespread, particularly in the non-alcoholic beverages industry, construction industry, along with science and education sector and political institutions, indicating that these industries may benefit from leveraging diverse references to enhance credibility and appeal to a more discerning audience. The “Direct User of the Product or Service” category of statements in analysed native articles is used in energetics, technological, financial, and food industries, reflecting the importance of authentic, relatable endorsements in areas where personal experience is particularly valued by consumers, but this category is generally used in lesser amount within different industries compared to other industries.

Industries such as alcoholic beverages and arts & culture occasionally employ sources unrelated to the product or service, which is the technique also used in betting & lottery sector and postal services, most probably to introduce a fresh perspective or to lend authority from a seemingly impartial standpoint. This technique is less common but can be effective in adding an element of unexpected credibility.

The “Influencer/Celebrity” category is notably prominent in sports industry, and is also used in media, marketing and digital agencies, along with furniture and home appliances industry, implying that these sectors rely on aspirational appeal and the social influence of well-known figures to drive engagement. Similarly to previous category, the use of brand ambassadors shows strength in sports industry, where a consistent and trustworthy spokesperson can help bridge the gap between complex products and the consumer, but this technique is also used in arts & culture and science & education sectors.

# Conclusion

The theoretical basis of this paper was the thesis that native advertising is “an extension of content marketing and provides marketers with a platform to place content in front of a wider audience they wouldn't ordinarily reach” [@crook2022]. Although there are authors who highlight certain criteria by which content marketing differs from native advertising, the ultimate goal is common: to drive profitable customer action. Based on the findings of this research, several conclusions can be drawn. The analysis of 543 native ads published on six Croatian online portals reveals a predominant use of clickbait headlines, with nearly 80% of headlines containing at least one clickbait characteristic. The most frequently employed technique is the style of uncertainty, which significantly contributes to audience engagement, as evidenced by its high average and total reach. This indicates that uncertainty-driven headlines are particularly effective in attracting readers. This is evidence that advertisers, through their content, aim to promote a “culture of fear”, a concept typically explored by sociologists [@maskalan2014; @furedi1997] in the context of political actions, but which is also applicable in the advertising industry. This is particularly due to the fact that fear/insecurity appeals can encourage people to change their attitudes and purchasing preferences [@foxall2007; @kesic2003]. Martin Lindstrom [@lindstrom2008] anticipated that contemporary advertising would increasingly be based on somatic markers triggered by fear and a sense of insecurity. In other words, advertisers will try to convince us that by not purchasing their product, we become insecure, dissatisfied, and unfree, which is evident in this analysis as well.

Interestingly, a smaller portion of the headlines (5.52%) mentioned the brand, suggesting that most ads are neutral in terms of direct brand promotion. This is, of course, in line with the habitus of native advertising, which is subtle, unobtrusive, and does not immediately reveal that it is, in fact, advertising. All analysed ads included visual elements, with photographs being the most common, either alone or combined with other types of visual content. The presence of the advertiser's logo or elements of their visual identity in 34.44% of the ads further underscores the importance of visual branding in native advertising.

Regarding the content characteristics of native ads, the analysis also considered the types of sources or statements used. The data show that almost half of the ads did not include any statements or sources, which is particularly common in certain industries such as retail, utility services, and shopping centres. When sources were used, the most frequent were individuals from the organization or company, indicating the value placed on insider credibility. Multiple sources and unrelated sources were less commonly used, but they demonstrated strong potential for high engagement, particularly when diverse perspectives were presented. The effectiveness of different source types varies across industries. For instance, the use of multiple sources is most effective in the non-alcoholic beverages and construction industries, where diverse references can enhance credibility. In contrast, direct user testimonials are more effective in sectors like technology and finance, where personal experience adds value to the advertising message. Clickbait techniques such as the use of numbers and vague pronouns also show varying levels of effectiveness depending on the industry. The retail and furniture industries, for example, favour the use of numbers to appeal to detail-oriented audiences. In contrast, emotional emphasis is more commonly used in sectors like utility services and education, where emotionally charged content can influence consumer behaviour.

Overall, this research highlights the strategic use of clickbait headlines and source types in native advertising as a content marketing strategy. The findings suggest that while certain techniques like uncertainty and multiple sources are particularly effective, the choice of strategy should be carefully tailored to the industry and target audience to maximize engagement. The data underscore the importance of understanding the specific characteristics and preferences of different sectors when designing native ads, as well as the significant role of visual content and source credibility in achieving advertising success.

# References